---
title: "Analysis of k"
output: html_notebook
author: Paulina Grunwald
---

Data source: https://www.gapminder.org/data/
Medical source references:
http://www.cvtoolbox.com/cvtoolbox1/cholesterol/ch.html


```{r}
#load packages
library("tidyr")
library("dplyr")
require("XLConnectJars")
require("XLConnect")
library(ggplot2)
library(reshape2)
library(datasets)
library(dplyr)
library(readxl)
library(countrycode)
library(gridExtra)
library(scales)

#set the directory where all the data files are placed
setwd("c:/Users/Paulina/Desktop/Udacity-Data-Analyst-Nanodegree/Project 4 - Explore and Sumarize data/Exercises/Food supply")

#print list of files in the chosen directory
list.files()

#IFC - Inficator food consumption (kilocalories/person&day)
#ITC - Mean total cholesterol index (mmol/L)
#Load ITC from excel file
ITC_male <- read_excel("ITC_male.xlsx", sheet = "Data")
ITC_female <- read_excel("ITC_female.xlsx", sheet = "Data")
IFC <- read_excel("food_consumption.xlsx", sheet = "Data")
```

The mean TC (Total Cholesterol) of the female/female population, counted in mmol per L; this mean is calculated as if each country has the same age composition as the world population.

If you total cholesterol is in the 5 to 6 mmol/L range, you have twice the risk of heart attack as compared to your risk if your level was well below 5 mmol/L.

Total cholesteror lowe limit: 3.0 - 3.6
Total cholesteror upper limit: 5.0-6.5

```{r}
#create A Data Frame Tbl
ITC_male <- tbl_df(ITC_male)
ITC_female <- tbl_df(ITC_female)
IFC <- tbl_df(IFC)

```


```{r}
#Change the name of the first column to "country"
colnames(ITC_female)[1] <- "country"
colnames(ITC_male)[1] <- "country"
colnames(IFC)[1] <- "country"
```

```{r}
dim(ITC_female)
```
```{r}
dim(ITC_male)
```

```{r}
dim(IFC)
```


```{r}
IFC
```


```{r}
df1 <- ITC_female %>%
  gather("year", "ITC", 2:30, convert = TRUE) %>%
  filter(!is.na(ITC)) %>%
  arrange(country,year, ITC) 

df2 <- ITC_male %>%
  gather("year", "ITC", 2:30, convert = TRUE) %>%
  filter(!is.na(ITC)) %>%
  arrange(country,year, ITC)

df3 <- IFC %>%
  gather("year", "IFC", 2:48, convert = TRUE) %>%
  filter(!is.na(IFC))  %>%
  arrange(country,year, IFC)
```

```{r}
unique(df3$year)
```

```{r}
unique(df1$year)
```


```{r}
colnames(df1)[3] <- "ITC_F"
colnames(df2)[3] <- "ITC_M"
```


```{r}
df <- left_join(df1, df2, by=c("country", "year"))
```


```{r}
df_final <- left_join(df,df3, by=c("year", "country")) %>%
  filter(!is.na(IFC)) 
```

```{r}
df_final$continent <- countrycode(df_final$country, "country.name", "continent")
```


```{r}
df_final
```

```{r}
unique(df_final$year)
```

```{r}
unique(df_final$continent)
```

Reordering the collumns in data frame
```{r}
df_final = df_final[c("country", "continent", "year", "IFC", "ITC_F", "ITC_M" )]
```

```{r}
df_final
```


```{r}
unique(df_final$year)
```

Peeking into the extremes of the data

```{r}
data2007 = subset(df_final, year == 2007)
summary(data2007)
```


```{r}
data1980 <- subset(df_final, year==1980)
  
summary(data1980)
```


```{r}
data2007 <- df_final %>%
  group_by(year==2007) 
  
summary(data2007)
```

```{r}
test <- df_final %>%
  group_by(country) %>%
  group_by(continent) %>%
  filter(year==2007) %>%
  summarise(medIFC = median(IFC))
test
```

```{r}
qplot(x = continent, y = ITC_M, geom = 'boxplot', data = data1980)+
  facet_wrap(~year, ncol = 1)
```


```{r}
ggplot(data = df_final, aes(year, y = IFC)) +
  geom_boxplot() + ylab("Food consumption indicator (kcal/day)")+
  theme(axis.text.x = element_text(angle=90, vjust = 0.5) )+
  facet_grid(~continent)
```


```{r}
ggplot(data = data1980, aes(x = continent, y = IFC)) +
  geom_boxplot() + 
  ylab("Food consumption indicator (kcal/day)") 
    
  
```

```{r}
ggplot(data = data2007, aes(x = continent, y = IFC)) +
  geom_boxplot() + 
  ylab("Food consumption indicator (kcal/day)") 
```


