Lesson 3
========================================================

***

### What to Do First?
Notes:
What questions are you trying to mention? What is your goal of exploratory data analysis?
You need to unserstand the most important and central to your analysis. Use summary and visualisation to explore the dat.
***

### Pseudo-Facebook User Data
Notes:
getwd() -> get the directory
list.files() -> get the list file in the directory
variable <- read.csv(' ', sep = '\t') -> load csv file with tab separator
```{r}
getwd()
```
```{r}
list.files()
```

The columns in the file are delimitated by tab this we need to use separator = '\t'

```{r Pseudo-Facebook User Data}
pf <- read.delim('pseudo_facebook.tsv',sep = '\t') 
```

```{r}
dim(pf)
```


You can see names of the variables using:
```{r}
names(pf)
```


***

### Histogram of Users' Birthdays

Notes:


```{r}
names(pf)
```

```{r}
install.packages('ggplot2')
library(ggplot2)

qplot(x = dob_day, data = pf, binwidth = 1)
```



```{r}

qplot(x = dob_day, data = pf) +
  scale_x_continuous(breaks=1:31)

```


***

#### What are some things that you notice about this histogram?
Response: Lot's of people have birthday on 1st of the month.

***

### Moira's Investigation
Notes:
Moira made an investigation to see if peoples perception of their audience matches the reality.

***

### Estimating Your Audience Size
Notes: People dramatically underestimate the size of their audience on facebook.


***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response: Message on that i was going on holidays to Mallorca

#### How many of your friends do you think saw that post?
Response: 30, 20%

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is? 
Response:
30%
***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:
facet_wrap(formula)
facet_wrap(~variable)


facet grid(formula)
facet_grid(vertical~horizontal)

```{r Faceting}
library(ggplot2)
qplot(x=dob_day, data = pf) +
  scale_x_continuous(breaks = 1:31) + 
  facet_wrap(~dob_month, ncol = 4)

```


```{r}
ggplot(data = pf, aes(x = dob_day)) + 
  geom_histogram(binwidth = 1) + 
  scale_x_continuous(breaks = 1:31) + 
  facet_wrap(~dob_month)
```

```{r}
ggplot(data = pf, aes(x = dob_day)) + 
  geom_histogram(binwidth = 1) + 
   scale_x_continuous(breaks = 1:31) + 
   facet_wrap(~dob_month)
```


#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}

library(ggplot2)
diamondsqplot(x=friend_count, data = pf, geom="histogram", binwidth = 10,  xlim=c(1,1200), ylim = c(0,600))


```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
qplot(x = friend_count, data = pf, xlim = c(0,1000))
```


```{r}
qplot(x = friend_count, data = pf) +
  scale_x_continuous(limits = c(0,1000)) +
  facet_wrap(~gender)
```



### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf, !is.na(gender)),
      binwidth = 10)+
  scale_x_continuous(limits = c(0,1000), breaks = seq(0,1000,100))+
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}

qplot(x=friend_count, data = pf, binwidth = 25)+ 
  scale_x_continuous(limits = c(0,1000), breaks = seq(0,1000,50))
```

#### Who on average has more friends: men or women?
Response:

```{r}
table(pf$gender)
```
It looks like men have more friends than woman

```{r}
by(pf$friend_count, pf$gender, summary)
```

```{r}
x = 96-74
x
```


#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes:

Examine the distribution of tenure or how many days someone has been using Facebook. 
Plots will include colour.
30 days is about motnh
```{r Tenure}
qplot(x = tenure, data = pf, bin = 30,
      color = I('black'), fill = I('#094DD8'))
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
tenuredf <- data.frame(pf$tenure/365)
                 
qplot(x = tenuredf, data = pf, binwidth = 1,
      color = I('black'), fill = I('#094DD8'))
                       
```


Solution from the course:

```{r}
qplot(x = tenure/365, data = pf, binwidth = 0.25,
      color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(breaks = seq(1,7,1), limits = c(0,7))
```

***

### Labeling Plots
Notes: We should adjust lables otherwise they will be automatically generated

```{r Labeling Plots}
qplot(x = tenure/365, data = pf, binwidth = 0.25,
      xlab = 'Number of years using Facebook',
      ylab = 'Number of users in sample',
      color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(breaks = seq(0,7,1), limits = c(0,7))
```

***

### User Ages
Notes:
Creates statistical summary of age data:
```{r}
summary(pf$age)
```

```{r User Ages}
qplot(x = age, data = pf, binwidth = 1,
      xlab = 'Age',
      ylab = 'Frequency',
      color = I('black'), fill = I('#F88725')) +
  scale_x_continuous(breaks = seq(13,110,5))
```

#### What do you notice?
Response:


```{r}
summary(pf$age)
```

***

### The Spread of Memes
Notes:



***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:
sometimes data are over dispersed. 

```{r}
p1 <- qplot(x = friend_count , data =pf,color = I("black"), fill = I("#F79420"))
p2 <- qplot(x = log10(friend_count + 1) , data =pf, color = I("black"), fill = I("#F79420"))
p3 <- qplot(x = sqrt(friend_count),  data =pf, color = I("black"), fill = I("#F79420"))

grid.arrange(p1, p2, p3, ncol = 1)

```





***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
library(gridExtra)
logScale <- qplot(x = log10(friend_count), data = pf)

countScale <- ggplot(aes(x= friend_count), data = pf) +
  geom_histogram()+
  scale_x_log10()

grid.arrange(logScale, countScale, ncol=2)

```

***


### Frequency Polygons

```{r Frequency Polygons}
qplot(x = friend_count, data = subset(pf, !is.na(gender)),
      binwidth = 10)+
  scale_x_continuous(lim = c(0,1000), breaks = seq(0,1000,50))

```


```{r}
```{r Frequency Polygons}
qplot(x = friend_count, data = subset(pf, !is.na(gender)),
      binwidth = 10, geom = 'freqpoly', color = gender) +
  scale_x_continuous(lim = c(0,1000), breaks = seq(0,1000,50))
 
```

```{r}
qplot(x = friend_count, y = ..count../sum(..count..), 
      data = subset(pf, !is.na(gender)),
      xlab = 'Friend Count',
      ylab = 'Proportion of Users with that friend count',
      binwidth = 10, geom = 'freqpoly', color = gender) +
  scale_x_continuous(lim = c(0,1000), breaks = seq(0,1000,100))
```



***

### Likes on the Web
Notes:
similar to histograms. but they draw curve connecting counts in histogram.

```{r Likes on the Web}

qplot(x = likes, y = ..count../sum(..count..), 
      data = subset(pf, !is.na(gender)),
      xlab = 'Likes Count',
      ylab = 'Proportion of Users with that friend count',
      binwidth = 10, geom = 'freqpoly', color = gender) +
  scale_x_continuous(lim = c(0,2000), breaks = seq(0,2000,100))

```


```{r}

qplot(x = www_likes, y = ..count../sum(..count..), 
      data = subset(pf, !is.na(gender)),
      xlab = 'Likes Count',
      ylab = 'Proportion of Users with that friend count',
      binwidth = 10, geom = 'freqpoly', color = gender) +
  scale_x_continuous(lim = c(0,1500), breaks = seq(0,1500,100))
```

```{r}
qplot(x = www_likes, data = subset(pf, !is.na(gender)),
      xlab = 'Likes Count',
      ylab = 'Proportion of Users with that friend count',
      geom = 'freqpoly', color = gender) +
  scale_x_continuous()+
  scale_x_log10()
```

Likes on the web solution
```{r}
by(pf$www_likes, pf$gender, sum)
```


***

### Box Plots
Notes:

```{r Box Plots}

qplot(x = gender , y = friend_count,
      data = subset(pf, !is.na(gender)), 
      ylim = c(0,1000),
      geom = 'boxplot')

```
Als we can use 

scale_y_continous(limits = c(0,1000)

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}

```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}

```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response: we will use friendship_initiated
```{r Friend Requests by Gender}

qplot(x = gender, y = friendships_initiated, data = subset(pf, !is.na(gender)), geom = 'boxplot') +
  coord_cartesian(ylim = c(0,250))



```
```{r}
by(pf$friendships_initiated, pf$gender, summary)
```

Response:

***

### Getting Logical
Notes:

The median in the first summary is 4 so it means we have a lot of 0s.

```{r Getting Logical}
summary(pf$mobile_likes)
```
```{r}
summary(pf$mobile_likes >0)
```

```{r}
mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in) #change to factor variable
summary(pf$mobile_check_in)
```

What procent of check in is using mobile?

Response:

```{r}

sum(pf$mobile_check_in == 1)/length(pf$mobile_check_in)
```
```{r}

***

### Analyzing One Variable
Reflection:

- always keep in mind what question are you trying to answer
- to get directory: getwd()
- to list files in the directory: list.files()
- to load csv we will use name <-read.csv('pseudo_facebook.tsv, sep = '\t')
- facet_wrap: splits data over a chosen variable
***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!