Lesson 5
========================================================

### Multivariate Data
Notes: 
We will analyze 3 or more variables at the time.

***

### Moira Perceived Audience Size Colored by Age
Notes:
http://hci.stanford.edu/publications/2013/invisibleaudience/invisibleaudience.pdf

When there is too much overplotting on the graph adding color as a new variable will most likely not help.

***

### Third Qualitative Variable
Notes:

```{r}
#load ggplot library
library(ggplot2)
library(dplyr)
library(tibble)
library(tidyr)
#load the pseudo facebook dataset
pf <- tbl_df(read.csv("pseudo_facebook.tsv", sep = '\t'))
```

```{r}
pf
```


```{r}
# Write code to create a new data frame,
# called 'pf.fc_by_age_gender', that contains
# information on each age AND gender group.

# The data frame should contain the following variables:

#    mean_friend_count,
#    median_friend_count,
#    n (the number of users in each age and gender grouping)

# Here is an example of the structure of your data frame. Your
# data values will be different. Note that if you are grouping by
# more than one variable, you will probably need to call the
# ungroup() function. 

pf.fc_by_age_gender <- pf %>%
  group_by(gender, age) %>%
  summarise(mean_friend_count = mean(friend_count), median_friend_count = median(friend_count), n=n()) %>%
  arrange(age, gender, mean_friend_count, median_friend_count, n) %>%
  ungroup() %>%
  arrange(age)
  
pf.fc_by_age_gender
```


```{r Third Qualitative Variable}

ggplot(aes(x = gender, y = age), data = subset(pf, !is.na(gender))) + 
  geom_boxplot()+
  stat_summary(fun.y = mean,geom = 'point', share = 5)
 
```

Almost always median friend count for woman is bigger than for men but there are certain exceptions e.g for very old users.




***

### Plotting Conditional Summaries
Notes:

Almost always median friend count for woman is bigger than for men but there are certain exceptions e.g for very old users.


```{r Plotting Conditional Summaries}
ggplot(aes(y = friend_count, x = age), data = subset(pf, !is.na(gender))) + 
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```

***

### Thinking in Ratios
Notes:
The plot can be useful if we want to investigate how median friend count varies with the age of the person.

***

### Wide and Long Format
Notes:
https://www.r-bloggers.com/melt/
http://seananderson.ca/2013/10/19/reshape.html
http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)
```
```{r}
#we use dcast because we want dataframe
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender,
                                  value.var = 'median_friend_count')
head(pf.fc_by_age_gender.wide)
pf.fc_by_age_gender
```


```{r}
#same code can be written using dplyr package
pf.fc_by_age_gender.wide2 <-     subset(pf.fc_by_age_gender[c('age', 'gender', 'median_friend_count')],                   !is.na(gender)) %>% 
    spread(gender, median_friend_count) %>% 
    mutate(ratio = male / female) 

head(pf.fc_by_age_gender.wide2)

```


***

### Ratio Plot
Notes:
http://docs.ggplot2.org/0.9.3.1/geom_hline.html

geom_hline(mapping = NULL, data = NULL, stat = "hline", position = "identity", show_guide = FALSE, 
  ...)

To create horizontal line y = 1 we need to use: geom_hline(mapping = aes(yintercept = 1))

```{r Ratio Plot}
# Plot the ratio of the female to male median
# friend counts using the data frame
# pf.fc_by_age_gender.wide.

# Think about what geom you should use.
# Add a horizontal line to the plot with
# a y intercept of 1, which will be the
# base line. Look up the documentation
# for geom_hline to do that. Use the parameter
# linetype in geom_hline to make the
# line dashed.

# The linetype parameter can take the values 0-6:
# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash

# This assignment is not graded and
# will be marked as correct when you submit.

ggplot(aes(y = ratio, x = age), data = pf.fc_by_age_gender.wide2)+
  geom_line()+
  geom_hline(mapping = aes(yintercept = 1),linetype = 4, alpha = 1 )

```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
# Create a variable called year_joined
# in the pf data frame using the variable
# tenure and 2014 as the reference year.

# The variable year joined should contain the year
# that a user joined facebook.

# See the Instructor Notes for three hints if you get
# stuck. Scroll down slowly to see one hint at a time
# if you would like some guidance.
pf$year_joined <- floor(2014 - pf$tenure/365)
pf
```

For positive numbers, trunc and floor give the same result. But for negative numbers, floor rounds down and trunc rounds up. This is because trunc always rounds toward 0.

```{r}
#Example of use floor() and trunc() function:
x <- c(-3.2, -1.8, 2.3, 2.9)
floor(x)
trunc(x)

```


***

### Cut a Variable
Notes:

https://www.r-bloggers.com/r-function-of-the-day-cut-2/
Cut() is useful to break up a continuous variable such as age into a categorical variable.

```{r Cut a Variable}
summary(pf$year_joined)
```

```{r}
# Create a new variable in the data frame
# called year_joined.bucket by using
# the cut function on the variable year_joined.

# You need to create the following buckets for the
# new variable, year_joined.bucket

#        (2004, 2009]
#        (2009, 2011]
#        (2011, 2012]
#        (2012, 2014]

# Note that a parenthesis means exclude the year and a
# bracket means include the year.

pf$year_joined.bucket <- cut(pf$year_joined,  c(2004,2009,2011,2012,2014))
table(pf$year_joined.bucket)

```


***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
table(pf$year_joined.bucket, useNA = 'ifany')
```

```{r Plotting it All Together}
ggplot(aes(x = age, y = friend_count), 
              data = subset(pf, !is.na(gender))) + 
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```


```{r}
# Create a line graph of friend_count vs. age
# so that each year_joined.bucket is a line
# tracking the median user friend_count across
# age. This means you should have four different
# lines on your plot.

# You should subset the data to exclude the users
# whose year_joined.bucket is NA.

ggplot(aes(x = age, y = friend_count), 
              data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```


***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
# Write code to do the following:

# (1) Add another geom_line to code below
# to plot the grand mean of the friend count vs age.

# (2) Exclude any users whose year_joined.bucket is NA.

# (3) Use a different line type for the grand mean.

# As a reminder, the parameter linetype can take the values 0-6:

ggplot(aes(x = age, y = friend_count), 
              data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)+
   geom_line(stat = 'summary', fun.y = mean, linetype = 2, color = 'blue')
```

***

### Friending Rate
Notes:

```{r Friending Rate}
with(subset(pf, tenure >= 1), summary(friend_count / tenure))
```


***

### Friendships Initiated
Notes:

What is the median friend rate?



What is the maximum friend rate?

```{r Friendships Initiated}
# Create a line graph of mean of friendships_initiated per day (of tenure)
# vs. tenure colored by year_joined.bucket.

# You need to make use of the variables tenure,
# friendships_initiated, and year_joined.bucket.

# You also need to subset the data to only consider user with at least
# one day of tenure.

ggplot(aes(x=tenure,y=friendships_initiated/tenure),data = subset(pf, tenure>=1))+
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)
 

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```


```{r}
ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)+
  geom_smooth(method = "lm", formula = y ~ x, size = 2)
```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv("yogurt.csv")
# Change the ID from an int to a factor
yo$id <-factor(yo$id)
str(yo)

```

```{r}
ggplot(aes(x=price, ), data = yo)+
  geom_histogram()
```

The data seem to be discrete as many prices are missing e.g we do not have any yogurts that would cost 30.

```{r}
ggplot(aes(x=price, ), data = yo)+
  geom_histogram(binwidth = 10)
```

If we set binsize to 10 we will not be able to observe this discretness. 

***

### Number of Purchases
Notes:

One clue to the discretness is the fact that 75% is very close to max. value.
```{r}
length(unique(yo$price))
```


```{r}
 table(yo$price)
```

```{r Number of Purchases}
# Create a new variable called all.purchases,
# which gives the total counts of yogurt for
# each observation or household.

# One way to do this is using the transform
# function. You can look up the function transform
# and run the examples of code at the bottom of the
# documentation to figure out what it does.

# The transform function produces a data frame
# so if you use it then save the result to 'yo'!

# OR you can figure out another way to create the
# variable

yo <- transform(yo, all.purchases=strawberry + blueberry + pina.colada + plain + mixed.berry)
all.purchases
```
```{r}
summary(yo$all.purchases)
```

```{r}
# alternate verbose from
yo$all.purchases <- yo$strawberry + yo$blueberry + yo$pina.colada + yo$plain + yo$mixed.berry
```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(aes(x=all.purchases), data=yo) +
  geom_histogram(fill = 'blue', binwidth = 1)
  
```

```{r}
ggplot(aes(x = time, y = price), data =yo)+
  geom_jitter(alpha=1/4, shape=21, fill = 'red')
 
```

Most common prices are increasing over the time. The lower scatter prices could be brought down due to cupons or other promotions.

***

### Sampling Observations
Notes:
If the dataset contains multiple obesrvation of the same unit it is often useful to work with the sample of those units. Then it is easier to display raw data for this sample. It will allow to bette understand within or in between household variation.


***

### Looking at Samples of Households

https://www.stat.berkeley.edu/classes/s133/factors.html
http://www.statmethods.net/advgraphs/parameters.html

* Use the pch or shape parameter to specify the symbol when plotting points.

```{r Looking at Sample of Households}
# Set the seed for reproducible results
set.seed(4230)
sample.ids <-sample(levels(yo$id), 16)

ggplot(aes(x=time, y=price), data = subset(yo, id %in% sample.ids)) +
  facet_wrap(~ id)+
  geom_line()+
  geom_point(aes(size = all.purchases), pch = 1)
```

```{r}
set.seed(3434)
sample.ids <-sample(levels(yo$id), 16)

ggplot(aes(x=time, y=price), data = subset(yo, id %in% sample.ids)) +
  facet_wrap(~ id)+
  geom_line()+
  geom_point(aes(size = all.purchases), pch = 1)+
  scale_x_continuous(breaks = seq(9800,10400,300))
  
```

We can observe that not in all household the yoghurt are bought regularly. There is variety of different purchasing patterns. The household 212305 has regular pattern of buying yogurts but at the times price is drawn lower - this could be caused by discount or coupons. Some households buy yoghurts very rarely e.g 2107706.

***

### The Limits of Cross Sectional Data
Notes:
We can not use time plot for facebook data.

***

### Many Variables
Notes:


***

### Scatterplot Matrix

Notes:
Scatterplot Matrix can be useful in the beginning of the analysis.

https://www.r-bloggers.com/scatterplot-matrices/

Scatterplot matrices are a great way to roughly determine if you have a linear correlation between multiple variables. This is particularly helpful in pinpointing specific variables that might have similar correlations to your genomic or proteomic data. If you already have data with multiple variables, load it up as described here.


ttps://s3.amazonaws.com/udacity-hosted-downloads/ud651/scatterplotMatrix.pdf
```{r}
library(GGally)

theme_set(theme_minimal(20))

# set the seed for reproducible results
set.seed(1836)
pf_subset <- pf[, c(2:15)]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset),1000),])
```




***

### Even More Variables
Notes:
```{r}
nci <- read.table('nci.tsv')

#changing the col names to produce a nicer plot
colnames(nci) <-c(1:64)
```


***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
#melt to a long format
library(reshape2)
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

