Lesson 3
========================================================

***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes:
```{r}
getwd()
```
```{r}
list.files()
```

The columns in the file are delimitated by tab this we need to use separator = '\t'

```{r Pseudo-Facebook User Data}
pf <- read.delim('pseudo_facebook.tsv',sep = '\t') 
```

```{r}
dim(pf)
```


You can see names of the variables using:
```{r}
names(pf)
```


***

### Histogram of Users' Birthdays

Notes:


```{r}
install.packages('ggplot2')
library(ggplot2)

names(pf)
qplot(x = dob_day, data = pf)
```



```{r}
install.packages('ggplot2')
library(ggplot2)

names(pf)
qplot(x = dob_day, data = pf) +
  scale_x_continuous(breaks=1:31)

```


***

#### What are some things that you notice about this histogram?
Response: Lot's of people have birthday on 1st of the month

***

### Moira's Investigation
Notes:
Moira made an investigation to see if peoples perception of their audience matches the reality.

***

### Estimating Your Audience Size
Notes: People dramatically underestimate the size of their audience on facebook.


***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response: Message on that i was going on holidays to Mallorca

#### How many of your friends do you think saw that post?
Response: 30, 20%

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is? 30%
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:
facet_wrap(formula)
facet_wrap(~variable)


facet grid(formula)
facet_grid(vertical~horizontal)

```{r Faceting}
library(ggplot2)
qplot(x=dob_day, data = pf) +
  scale_x_continuous(breaks = 1:31) + 
  facet_wrap(~dob_month, ncol = 3)

```


```{r}
ggplot(data = pf, aes(x = dob_day)) + 
  geom_histogram(binwidth = 1) + 
  scale_x_continuous(breaks = 1:31) + 
  facet_wrap(~dob_month)
```

```{r}
ggplot(data = pf, aes(x = dob_day)) + 
  geom_histogram(binwidth = 1) + 
   scale_x_continuous(breaks = 1:31) + 
   facet_wrap(~dob_month)
```


#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}

library(ggplot2)
qplot(x=friend_count, data = pf, geom="histogram", binwidth = 10,  xlim=c(1,1200), ylim = c(0,600))


```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
qplot(x = friend_count, data = pf, xlim = c(0,1000))
```


```{r}
qplot(x = friend_count, data = pf) +
  scale_x_continuous(limits = c(0,1000)) +
  facet_wrap(~gender)
```



### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf, !is.na(gender)),
      binwidth = 10)+
  scale_x_continuous(limits = c(0,1000), breaks = seq(0,1000,100))+
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}

qplot(x=friend_count, data = pf, binwidth = 25)+ 
  scale_x_continuous(limits = c(0,1000), breaks = seq(0,1000,50))
```

#### Who on average has more friends: men or women?
Response:

```{r}
table(pf$gender)
```
It looks like men have more friends than woman

```{r}
by(pf$friend_count, pf$gender, summary)
```

```{r}
x = 96-74
x
```


#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes:

Examine the distribution of tenure or how many days someone has been using Facebook. 
Plots will include colour.

```{r Tenure}
qplot(x = tenure, data = pf, bin = 30,
      color = I('black'), fill = I('#094DD8'))
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
tenuredf <- data.frame(pf$tenure/365)
                 
qplot(x = tenuredf, data = pf, binwidth = 1,
      color = I('black'), fill = I('#094DD8'))
                       
```


Solution from the course:

```{r}
qplot(x = tenure/365, data = pf, binwidth = 0.25,
      color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(breaks = seq(1,7,1), limits = c(0,7))
```

***

### Labeling Plots
Notes: We should adjust lables otherwise they will be automatically generated

```{r Labeling Plots}
qplot(x = tenure/365, data = pf, binwidth = 0.25,
      xlab = 'Number of years using Facebook',
      ylab = 'Number of users in sample',
      color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(breaks = seq(0,7,1), limits = c(0,7))
```

***

### User Ages
Notes:

```{r User Ages}
qplot(x = age, data = pf, binwidth = 1,
      xlab = 'Age',
      ylab = 'Frequency',
      color = I('black'), fill = I('#F99425')) +
  scale_x_continuous(breaks = seq(0,110,5))
```

#### What do you notice?
Response:


```{r}
summary(pf$age)
```

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:


```{r}
p1 <- qplot(x = friend_count , data =pf,color = I("black"), fill = I("#F79420"))
p2 <- qplot(x = log10(friend_count + 1) , data =pf, color = I("black"), fill = I("#F79420"))
p3 <- qplot(x = sqrt(friend_count),  data =pf, color = I("black"), fill = I("#F79420"))

grid.arrange(p1, p2, p3, ncol = 1)

```





***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}

```

***


### Frequency Polygons

```{r Frequency Polygons}

```

***

### Likes on the Web
Notes:

```{r Likes on the Web}

```


***

### Box Plots
Notes:

```{r Box Plots}

```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}

```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}

```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}

```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}

```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!