---
title: "Oil consumption per capita evaluation"
author: Paulina Grunwald 
output: html_notebook
---

Source:- http://www.gapminder.org/data/

```{r}
install.packages("tidyr")
```


```{r}
library("tidyr")
library("dplyr")
library(ggplot2)
library(reshape2)
library(datasets)
library(dplyr)
require(xlsx)
library(countrycode)
library(gridExtra)
library(scales)

setwd("c:/Users/Paulina/Desktop/Udacity-Data-Analyst-Nanodegree/Project 4 - Explore and Sumarize data/Exercises")
list.files()
getwd()

#OCPC - Oil consumption per capita
OCPC <- tbl_df(read.csv("OilConsumptionPerCapita.csv", header=TRUE, check.names = FALSE))
#HDI - human development index
IHDI <- read.csv('Indicator_HDI.csv', header=TRUE, check.names = FALSE)

```


```{r}
OCPC
```

```{r}
IHDI
```


Now i will change first column name:

```{r}
colnames(IHDI)[1] <- "country"
colnames(OCPC)[1] <- "country"
```

We will use gather() function in order to change layout of the dataframe (from wide data frame format into long one). Gather() breakdowns numerous columns into two key-value pairs. It's an alternative of melt() function [in reshape2 package]. 

Format of the gather() function is following: gather(data, key, value, ...)
* data: A data frame
* key, value: Names of key and value columns to create in output
* . : Specification of columns to gather. Allowed values are:
  - variable names
  - if you want to select all variables between a and e, use a:e
  - if you want to exclude a column name y use -y
  - for more options, see: dplyr::select()

```{r}
#dimension of the data frame
dim(OCPC)
```
```{r}
dim(IHDI)
```


convert = TRUE converts numbers in the keys column from factor to numerics.
```{r}
OCPC2 <- OCPC %>%
  gather("year", "OCPC", 2:47, convert = TRUE) %>%
  arrange(country, year, OCPC)
OCPC2
```


```{r}
IHDI2 <- IHDI %>%
  gather("year", "HDI", -1, convert = TRUE) %>%
  arrange(country, year, HDI)
IHDI2
```



```{r}
df <- OCPC2


df = na.omit(df)
df = data.frame(df, row.names = NULL, check.names = FALSE)
df %>%
  arrange_("country", "year")
```

```{r}
com_df2 <- com_df %>% 
  filter(!is.na(OCPC), !is.na(HDI)) %>%
  arrange_("country","year")
com_df2
```
```{r}
qplot(x = year, data =df, y = OCPC)+
  scale_x_continuous(breaks = 1:10)+
  facet_wrap(~country, ncol = 5)

  
```

```{r}
qplot(x = year, data =df, y = OCPC)+
  scale_x_continuous(breaks = 1:10)+
  scale_y_continuous(break = 1)
  facet_wrap(~country)
```

```{r}
install.packages("ggvis")
library(ggvis)
```

```{r}
algeria <- filter(com_df2, country %in% c("Algeria"))
algeria %>%
  ggvis(x = ~year, y = ~OCPC, data = ~HDI)

```



```{r}
df2$continent <- countrycode(df2$country, "country.name", "continent")
head(df2)
```HDI

```{r}
by((subset(df2$!is.na(OCPC)), df2$country, median))
```

```{r}
mean(df2$OCPC,df2$country=='Poland', na.rm=T)
```

median()
sd()
quantile(, probs = c(0.01,0.05,0.25,0.5,0.75,0.9,0.95,0.99)

summary()

library(Hmisc)
describe()

pastecs::stat.desc()
#it gives coef of var, and CI mean 95%