Survival chances of the Titanic's passangers Analysis by Paulina Grunwald
============================================================================

---
date: "March 3, 2017"
output: html_document
---

<style type="text/css">

body, td {
   font-size: 18px;
}

</style>


```{r global_options, include=FALSE}

knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```



##Introduction
******
In this data visualization project in d3.js and dimple.js  I will use Titanic dataset. The dataset contains information of 891 passangers out of 2224 passangers and crew on board of the Titanic. The dataset contains information such as passangers' age, sex, embarkment location, in which class were they located, ticket price, cabin number etc. The purpose of my data visualization is primarily to show which passangers had higher chance to survive looking at the sex, class and age variables. My visualizaions show that female and children had higher chance of survival than men. The highest survival rate was observed in passangers form the first class. Most of the males from the third class passengers did not survive the disaster.


##Dataset wrangling
******
Titanic dataset was provided in csv format. As a start of the data preparation process I have loaded the titanic csv file into R studio in order to clean the dataset and prepare final dataframe that will be used for my visualization.

```{r echo=FALSE,warning=FALSE, message=FALSE}
#Load packages
require(knitr)
require(markdown)
library(tidyr)
library(dplyr)
library(ggplot2)
library(RColorBrewer)

#Load csv file
titanic <-  tbl_df(read.csv("C:/Personal/GitHub/Udacity-Data-Analyst-Nanodegree/Project 6 - Make Effective Data Visualization/Final project/Second Revision/titanic_data.csv"))
```

Below you can find part of the initital dataset:

```{r echo=FALSE,titanic_dataset}
#View titanic dataset
titanic
```


Since i have decided to use only age, sex and and class variables I will now remove other columns which i will not use in my investigation. I will also omit all passangers that have NA values for any of the variables. 

```{r echo=FALSE}
#Remove not needed columns
titanic$PassengerId <- NULL
titanic$Cabin <- NULL
titanic$Name <- NULL
titanic$Ticket <- NULL
titanic$Fare <- NULL
titanic$Cabin <- NULL
titanic$SibSp <- NULL
titanic$Parch <- NULL

#Omit NA values
titanic<- na.omit(titanic)
titanic[complete.cases(titanic),]
```

Now i will view summary of my cleaned dataset:

```{r echo=FALSE}
#Display statistical summary of titanic dataset
summary(titanic)
```

I will continue with cleaning process. I have decided to create new columns in order to get more insights from the dataset. 
```{r echo=FALSE}

#Create new column Age_Type to separate aduls and childen into two groups
titanic$Age_Type  = ""
titanic$Age_Type[titanic$Age < 18] = "Child"
titanic$Age_Type[titanic$Age >= 18] = "Adult"

#Assign 0 to Perished and 1 to survived in Survived column for easier understanding
titanic$Survived[titanic$Survived == 0] = "Perished"
titanic$Survived[titanic$Survived == 1] = "Survived"

#Rename Pclass column to Class
colnames(titanic)[colnames(titanic)=="Pclass"] <- "Class"

#Rename Class number of for easier understanding
titanic$Class[titanic$Class == 1] = "First"
titanic$Class[titanic$Class == 2] = "Second"
titanic$Class[titanic$Class == 3] = "Third"

#Rename Location of embarkment for easie understanding
titanic$Embarked <- as.character(titanic$Embarked)
titanic$Embarked[titanic$Embarked == "S"] = "Southampton"
titanic$Embarked[titanic$Embarked == "Q"] = "Queenstown"
titanic$Embarked[titanic$Embarked == "C"] = "Cherbourg"

#Create Age_Group column - passangers will be separated into age buckets
titanic$Age_Group   = ""
titanic$Age_Group <- as.character(titanic$Age_Group )
titanic$Age_Group[titanic$Age <= 10] = "0-10"
titanic$Age_Group[titanic$Age > 10 & titanic$Age <= 20] = "11-20"
titanic$Age_Group[titanic$Age > 20 & titanic$Age <= 30] = "21-30"
titanic$Age_Group[titanic$Age > 30 & titanic$Age <= 40] = "31-40"
titanic$Age_Group[titanic$Age > 40 & titanic$Age <= 50] = "41-50"
titanic$Age_Group[titanic$Age > 50 & titanic$Age <= 60] = "51-60"
titanic$Age_Group[titanic$Age > 60 & titanic$Age <= 70] = "61-70"
titanic$Age_Group[titanic$Age > 70 & titanic$Age <= 80] = "71-80"

#Create class_sex column to seperate the passangers by class sex
titanic$Class_Sex   = ""
titanic$Class_Sex <- as.character(titanic$Class_Sex )
titanic$Class_Sex[titanic$Class == "First" & titanic$Sex == "female"] = "1st Class Female"
titanic$Class_Sex[titanic$Class == "First" & titanic$Sex == "male"] = "1st Class Male"
titanic$Class_Sex[titanic$Class == "Second" & titanic$Sex == "female"] = "2nd Class Female"
titanic$Class_Sex[titanic$Class == "Second" & titanic$Sex == "male"] = "2nd Class Male"
titanic$Class_Sex[titanic$Class == "Third" & titanic$Sex == "female"] = "3rd Class Female"
titanic$Class_Sex[titanic$Class == "Third" & titanic$Sex == "male"] = "3rd Class Male"


#Create count column which will be filled with value of "1". This will help while creating barplot count in my d3.js visualization
titanic$Count = ""
titanic$Count <- as.character(titanic$Count)
titanic$Count = "1"

```

Final dataframe
```{r}
titanic
```

Once the data wrangling process was finished, I have  exported the cleaned and reshaped dataframe into csv file. This file will be used for creating the visualization in D3.js
S
```{r echo=FALSE, export_to_csv}
# Write CSV in R with headers
write.table(titanic, file = "titanic.csv",row.names=FALSE, na="",col.names=TRUE, sep=",")
```


##Design
******
In order to create the data visualization, I have started first with the investigation of the data. I have tried to choose the most important data categories associated with the Titanic passangers. In my opinion the most important are Sex, Class and Age. I believe that they had the biggest impact on the chance of survival of the passangers. As a prefered data visualization plot for my project i have chosen stacked bar plot because it provides a visual presentation of categorical data. I wanted to stay away from pie chart as they can be often confusing.

For bar color selection i have used colorbrewer2.org. I wanted to use colorblind safe color scheme. I have chosen qualitative Dark2 palette. To visualize survived passangers I have chosen green color (color of life) and for perished passangers red color (color of blood). The color and the order of axis are keep consistent over different charts to help the viewer to have clear understanding of the story that I am trying to convey.

I have decided to create bar plots including age, sex and class. I have found that those three variables and survival rate are very closely linked. The first plot represents survival rate by sex. I wanted to show that men had much small survival chance in comparision to woman. It is possible that woman had priority while entering the life boats. The second and third plot represents survival rate by age group. In the second plot I separated the passengers into adults and children (children were considered as <=18 years old). On the third plot I presented survival rate of passengers in 8 different age groups (0-10, 11-20, 21-30, 41-50, 51-60, 61-70, 71-80 ). I have also added the button that changes the y axis to procent and then back to count. I did that in order to see exact procentages of survived and persihed passengers in all age groups (for example it was difficoult to see exacly the ratio of survivors to perished in last two age groups). On the fourth plot i have explored survival rate by class. I wanted to show that passangers in the third class were the most affected by titanic tragedy. On the fifth plot i added another layer to the fourth plot which was sex data. I wanted to show viewers how many men and women survived in class one, two and three. 

##Feedback
******

After finishing my virst visualization (the visualization can be viewved at http://bl.ocks.org/paulina-grunwald/4914047cebb29ad0227fe086f164cf59) I have posted it in the Udacity website forum as well as Google+ Udacity group in order to get feedback from fellow students. I used the questions provided as guidance to ask for specific feedback such as:

What do you notice in the visualization?
What questions do you have about the data?
What relationships do you notice?
What do you think is the main takeaway from this visualization?
Is there something you don't understand in the graphic?
After submission of project for review I received feedback from the reviewer which is also added below. After my second submission, I got another feedback to do some finishing touches for polishing up the visualization.


I got the following feedback:

Feedback from Shahrzad (Udacity data analyst forum):

> Your Visualization looks good! I want to know which age group and which gender was more successful to get survived. In the left bar chart I can conclude 
> that female passengers had  priority to be survived rather than male passengers. In the right bar chart, I cannot say for sure that children or adults 
> had priority to be survived. If you are able to compare the proportions then we would be able to say which age group had higher survival rate. For 
> example if you compare the proportions of the survived people in each age group through the >stacked percent bar chart then I think we can make better 
> conclusions.

Feedback from xueming_663398266218 (Udacity data analyst forum):

> The visualisation is very clean with nice colour, the story you want to tell is also clear for me. Only one minor question: In the last figure, you used 
> PctAxis as the y axis, while you used counts for the other charts, is there >a special reason?

Feedback from Yannis Pappas (Google +)

> Nice visualization.  Some suggestion I could give is that you missed the title on your second bar plot (Age_Type) and that on the last two bar plots it is > not clear which column represent which gender.

Feedback Sarah Silva (Google +)

> I really appreciated the overall contextualization you gave, leveling everyone's knowledge about the story. The label "Survived: Perished" was a bit 
> confusion also. I found it interesting the conclusions about which sex and class had a higher survival rate. I don't know why the embarking location would > be important, though. That's all, congrats!

##Post-feedback
******
Following the feedback from the 4 different udacity students, I have implemented the changes in my visualization:

- I have added 2nd age group bar plot to show exacly people in which age group had the highest chance of survival,
- I have added buttons that change y Axis to procent and back to count for plot 3 (age group plot)
- I have added titles and corrected y and x axis names for all plots
- I have kept the "survived" and "perished" words in lengend of all of my graphs. I believe that they are self explenatory.
- I have removed plot with embarked location information as it didn't serve very well the story i was trying to convey.
- I have changed the colours of the plots to the ones that are more colorblind friendly (i have used Colorbrewer to find new colors).

The feedback from my fellow students helped me to improve my visualiztion and made me think about few other improvements that i have introduced in my final visualization.

Improved visualization can be accessed at: http://bl.ocks.org/paulina-grunwald/1b38df3ac5faa3ce0f632d4797ad6972


##Rescources:
******
https://github.com/PMSI-AlignAlytics/dimple/wiki/dimple.color
http://stackoverflow.com/questions/25416063/title-for-charts-and-axes-in-dimple-js-charts
https://bost.ocks.org/mike/bar/
http://dimplejs.org/examples_viewer.html?id=bars_vertical_grouped
https://en.wikipedia.org/wiki/RMS_Titanic
http://chimera.labs.oreilly.com/books/1230000000345
http://colorbrewer2.org/#type=qualitative&scheme=Dark2&n=3
https://en.wikipedia.org/wiki/Bar_chart